---
title: "project_tooth"
author: "lipnerova"
date: "December 21, 2015"
output: pdf_document
---

# Info on data:
The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).

# Zadání:
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

- Load the ToothGrowth data and perform some basic exploratory data analyses 
- Provide a basic summary of the data.
- Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
- State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on

- Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
- Did the student perform some relevant confidence intervals and/or tests?
- Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
- Did the student describe the assumptions needed for their conclusions?
    
## Loading the data

```{r libraries, message=FALSE, warning=FALSE}
# Load libraries
library(dplyr)
library(ggplot2)
library(gridExtra)
library(knitr)

# Load dataset
library(datasets)   
data(ToothGrowth)
a<-(ToothGrowth) #give it easily typed name
```


# EDA

```{r boxplots}
p1 <- ggplot(a, aes(y=len, x=as.factor(supp))) + 
  geom_boxplot(aes(fill=supp)) + 
  guides(fill=FALSE) +
  facet_grid(.~dose) +
  labs(title="Teeth growth on two supplements 
       \n by dose (mg/day) of supplement",
       x="Supplement type",
       y="Length of teeth") +
  theme(title = element_text(size = rel(0.75), hjust = 0.5))


p2 <- ggplot(a, aes(x=supp, y=len)) + geom_boxplot(aes(fill=supp)) + 
  labs(title="Overall teeth growth \n on two supplements",
       x="Supplement type",
       y="Length of teeth") +
  scale_fill_discrete(name="Supplement",
                      labels=c("Orange juice", "Vitamin C")) +
  theme(title = element_text(size = rel(0.75), hjust = 0.5))

# Arrange them: 
grid.arrange(p1, p2, ncol=2, nrow=1)
```

```{r tab_one}
a_tab<-group_by(a, supp, dose) %>% summarise(mean(len))
kable(x=a_tab, digits=1, align="c", caption="Summary table of data")
```


# Testing

Test hypothesis, that there is a difference between orange juice and vitamin C (all doses together.
Test each dose separately for above stated difference => together 4 tests.

## Hypotheses
Overall test: H0 there is no difference between OJ and VC
H1: the OJ has bigger effect than VC
```{r}
t.test(len ~ factor(supp), data=a, paired=F) 
```

PArtial tests - dose 0.5
```{r}
# Dose of 0.5 mg/day:
t.test(len ~ supp, data=a[a$dose==0.5,])

# Dose of 1 mg/day:
t.test(len ~ supp, data=a[a$dose==1,])

# Dose of 2 mg/day:
t.test(len ~ supp, data=a[a$dose==2,])


```

udělat tabulku s p hodnotami a CI