---
title: "project_tooth"
author: "lipnerova"
date: "December 21, 2015"
output: pdf_document
---

Zadání:
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

- Load the ToothGrowth data and perform some basic exploratory data analyses 
- Provide a basic summary of the data.
- Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
- State your conclusions and the assumptions needed for your conclusions. 

Some criteria that you will be evaluated on

- Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
- Did the student perform some relevant confidence intervals and/or tests?
- Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
- Did the student describe the assumptions needed for their conclusions?
    
## Loading the data

```{r libraries, message=FALSE, warning=FALSE}
# Load libraries
library(ggplot2)
library(dplyr)
library(knitr)

# Load dataset
library(datasets)   
data(ToothGrowth)
a<-(ToothGrowth) #give it easily typed name
```


# EDA

```{r scatterplot_one}
ggplot(a, aes(y=len, x=dose, group=supp)) + geom_point() + facet_grid(supp~.)
```

```{r tab_one}
a_tab<-group_by(a, supp, dose) %>% summarise(mean(len))
kable(x=a_tab, digits=1, align="c", caption="Summary table of data")
```

```{r}

```